Presentation Title

### 2017-08-22 thesis/chapters/ch4-analysis.tex
Edit grammar etc.

In this chapter,
we evaluate our system analytically to check whether it correctly implements the TrustChain consensus protocol (\Cref{def:consensus})
and the TrustChain validation protocol (\Cref{def:validation}).
Further, we analyse the performance and especially the throughput to understand the scalability characteristics.
Finally, we consider the case where the $n \ge 3t + 1$ assumption is violated and show the effect of it in our system.

\section{Correctness analysis}
\label{sec:correctness}
Our first objective is to show that our consensus protocol $\Pi_c$ described in~\Cref{sec:cons-protocol} is in fact a TrustChain consensus protocol defined in~\Cref{def:consensus}.
Then, building on top of it, we show that our validation protocol $\Pi_v$ described in~\Cref{sec:vd-protocol} conforms to the TrustChain validation protocol defined in~\Cref{def:validation}.
The resulting theorem shows that only using CP blocks in the consensus algorithm implies consensus on TX blocks.

@@

\begin{lemma}
\label{lemma:agreement}
If for an arbitrary round $r$,
$\F_{r}$ is known by all correct nodes and one correct node outputs a list of facilitators $\F_{r+1}$,
then all correct nodes output $\F_{r+1}$.
\end{lemma}
\begin{proof}
The argument follows from the protocol description.
Given that $\F_r$ is known,
correct nodes will send CP blocks to all members in $\F_r$.
The ACS algorithm starts independently whenever the facilitator has $N - t$ valid CP blocks
(recall from \Cref{sec:consensus-phase} that invalid blocks are ones with an invalid signature or has a duplicate signature).
It cannot make progress until $n-t$ honest facilitators start algorithm,
but this eventually happens because there are $N - t$ correct nodes and all correct facilitator eventually receives $N -t$ valid CP blocks.
At the end of ACS, $\C_{r+1}$ is created and then broadcasted along with the signature of the facilitators.
Due to the agreement property of ACS (\Cref{def:acs}),
every correct node should receive at least $n - t$ valid signatures on the agreed $\C_{r+1}$.
Thus they use $\C_{r+1}$ to generate a new CP block and compute new facilitators.
Since $\textsf{get\_facilitators}(\cdot)$ is a deterministic algorithm and the input $\C_{r+1}$ is in agreement, the output $\F_{r+1}$ is also in agreement.
\end{proof}

@@

\begin{lemma}
\label{lemma:fairness}
If for an arbitrary round $r$,
$\F_r$ is known by all correct nodes then
every node with a CP block in $\C_{r+1}$ should have an equal probability to be elected as a facilitator in $\F_{r+1}$.
\end{lemma}
\begin{proof}
We already established that $|\C_{r+1}| \ge N - t \ge n$ from \Cref{lemma:validity}.
Then the proof directly follows from the random oracle model.
Recall that the luck value is computed using $\textsf{H}(\C_{r+1} || pk_u)$.
Since $pk_u$ is unique for every node that has a CP block in $\C_{r+1}$,
the output of $\textsf{H}(\cdot)$ is uniformly random.
This effectively generates a random permutation,
so every node has the same probability of being in the top $n$ for the ordered sequence,
namely the output of $\textsf{get\_facilitators}(\C_{r+1})$.
\end{proof}

@@

$\Pi_v$ directly uses \Cref{alg:get-validity} which is also the validity definition,
so the validity property is also satisfied.
\end{proof}

\Cref{theorem:validation-agreement} shows that agreement on CP blocks would lead to agreement on TX blocks when the nodes are running the validation protocol.
Like in our prior work~\cite{implicitconsensus}, we call this behaviour implicit consensus.
One of the main advantages of our scheme over running a consensus algorithm on all the transactions is that 
the transaction throughput is no longer dependent on the consensus algorithm---ACS.
This enables horizontal scalability where adding new nodes would lead to higher global transaction rate.
In addition, a convenient consequence of \Cref{theorem:validation-agreement} is unforgeability.
That is, no adversary is able to create two chains,
$F = \{ \dots, t, c\}$ and $F' = \{ \dots, t', c\}$,
with correct hash pointers and the same end of chain $c$ with non-negligible probability.
% Another way to think about it is if a transaction can be forged, then it cannot be agreed.

@@

\begin{lemma}
There exists a valid transaction that cannot be eventually validated.
\end{lemma}
\begin{proof}
Suppose nodes $u$ and $v$ correctly performed the TX protocol to create a transaction.
Then when $u$ wants to validate it, it does so by sending \texttt{vd\_req} message to $v$.
$v$ can act maliciously and ignore all \texttt{vd\_req} message from $u$, then the transaction can never be validated.
\end{proof}
Although this is a negative result, it does not put the adversary in an advantageous position.
If the adversary is observed to ignore validation requests, then the honest nodes may prefer not to transact with her in the future.
Thus, to stay relevant in the system, the adversary needs to comply with the protocol.

@@

\section{Performance analysis}
\label{sec:performance-analysis}
This section aims to analytically answer whether our system is horizontally scalable.
We begin by looking at the communication and time complexity of the consensus protocol $\Pi_c$,
and then the bandwidth requirement for a single transaction.
We build on top of those results to analyse the global throughput.

@@

\subsection{Duration of the consensus protocol}
\label{sec:cons-duration}
In order to make arguments on the duration of $\Pi_c$, the bandwidth or throughput in our purely asynchronous model,
which are concepts that depend on time,
we must make additional assumptions.
Note that the duration we are interested in is not the same as the time complexity typically used in distributed systems.
In the analysis of distributed algorithms, time complexity is often in terms of the number of rounds.
For example, ACS runs in a constant number of rounds because of its sub-protocols---reliable broadcast and binary Byzantine consensus---also run in a constant number of rounds~\cite{miller2016honey}.
However, in practice, making a unit of communication always has some overhead associated with it, for example serialising and writing it to some network socket.
Hence, for the remainder of our performance analysis, we add the following to our computational model.
For every unit of communication, we assume they take some non-negligible but constant time to perform.
Hence, from~\Cref{sec:cons-complexity} and the fact that $\Pi_c$ runs in a constant number of rounds,
it follows that $\Pi_c$ has a duration of $O(N)$.

\subsection{Communication cost for transactions}
\label{sec:communication-cost-for-tx}
Geared with the results above,
we are ready to analyse the amount of data required to be transmitted over a link per transaction,
which we call the communication cost per transaction.
To create and then validate a transaction,
the communication cost per transaction is of $O(l)$,
where $l$ is the length of the agreed fragment.
This can be seen from the fact that the largest message by far is the \texttt{vd\_resp} message,
which contains the agreed fragment.
The other messages (\texttt{tx\_req}, \texttt{tx\_resp} and \texttt{vd\_req}) are constant factors.
If we assume that every node performs transactions at a constant rate of $r_{\text{tx}}$ per second.
Then
$$l = r_{\text{tx}} D_{\text{c}},$$
where $D_{\text{c}}$ is the duration of a round of the consensus protocol.
But from \Cref{sec:cons-duration},
we know that $D_{\text{c}}$ is of $O(N)$, thus the communication cost per transaction is $O(N)$.
This result is intuitive because round duration would be longer if there are more CP blocks (more $N$), which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
This behaviour is verified experimentally in \Cref{ch:implementation}.

\subsection{Linear global throughput}
\label{sec:global-throughput}
Using our results so far, we are able to analyse the global throughput.
First, we clarify the bandwidth definition, which is ``the data rate at which a network link or a network path can transfer''~\cite{prasad2003bandwidth}.
Now, suppose every node has $N$ links which connect to every other node,
every link has a fix bandwidth $C$ and every node makes $r_{\text{tx}}$ transactions per second.
Then we have the inequality
$$NC \ge r_{\text{tx}} l,$$
where $l$ is the length of the agreed fragment as before.
The inequality suggests that the rate for which transactions and validations are made cannot exceed the total bandwidth of all the links.

We note that the inequality does not hold if the node is only transacting with a subset of the population.
This is because it cannot use all the bandwidth available in all the links.
In the extreme case, if the node is only transacting with one other node, then it can only use the bandwidth of one link which is only $C$.
However, if that is the case, we can intelligently cache the \texttt{vd\_resp} messages as described in~\Cref{sec:caching}.
Hence, we analyse the worst case where every node transacts with a random node from the population,
and a new \texttt{vd\_resp} must be sent for every \texttt{vd\_req} message.

Consider the case where the system is making use of all the bandwidth, i.e. $NC = r_{\text{tx}} l$.
Recall that $l$ is of $O(N)$, that means LHS and RHS both grow linearly with respect to $N$.
Hence, there exists some $r_\text{tx}$ that makes use of all the available bandwidth regardless of $N$.
Finally, if every node in the network is transacting at $r_{\text{tx}}$ per second,
then the global throughput (in terms of transactions per second) is linear w.r.t. the population size $N$.
If the system is not making use of all the bandwidth.
We also maintain a constant transaction rate by the same argument,
so the global throughput is also linear w.r.t. $N$.
We verify both of these claims experimentally in~\Cref{ch:implementation}.

>>>>

In this chapter,
we evaluate our system analytically to check whether it correctly implements the TrustChain consensus protocol (\Cref{def:consensus})
and the TrustChain validation protocol (\Cref{def:validation}).
Further, we analyse the performance and in particular the throughput to understand the scalability characteristics.
Finally, we consider the case where the $n \ge 3t + 1$ assumption is violated and show the effect of it in our system.

\section{Correctness analysis}
\label{sec:correctness}
Our first objective is to show that our consensus protocol $\Pi_c$ described in~\Cref{sec:cons-protocol} is,
in fact, a TrustChain consensus protocol defined in~\Cref{def:consensus}.
Then, building on top of it, we show that our validation protocol $\Pi_v$ described in~\Cref{sec:vd-protocol} conforms to the TrustChain validation protocol defined in~\Cref{def:validation}.
The resulting theorem shows that only using CP blocks in the consensus algorithm implies consensus on TX blocks.

@@

\begin{lemma}
\label{lemma:agreement}
If for an arbitrary round $r$,
$\F_{r}$ is known by all correct nodes and one correct node outputs a list of facilitators $\F_{r+1}$,
then all correct nodes output $\F_{r+1}$.
\end{lemma}
\begin{proof}
The argument follows from the protocol description.
Given that $\F_r$ is known,
correct nodes will send CP blocks to all members of $\F_r$.
For each facilitator, the ACS algorithm starts independently whenever the facilitator has $N - t$ valid CP blocks
(recall from \Cref{sec:consensus-phase} that invalid blocks are ones with an invalid signature or has a duplicate signature).
No honest facilitator can make progress until $n-t$ honest facilitators start algorithm,
but this eventually happens because there are $N - t$ correct nodes and all correct facilitator eventually receive $N -t$ valid CP blocks.
At the end of ACS, $\C_{r+1}$ is created and then broadcasted along with the signature of the facilitators.
Due to the agreement property of ACS (\Cref{def:acs}),
every correct node should receive at least $n - t$ valid signatures on the agreed $\C_{r+1}$.
Thus they use $\C_{r+1}$ to generate a new CP block and compute new facilitators.
Since $\textsf{get\_facilitators}(\cdot)$ is a deterministic algorithm and the input $\C_{r+1}$ is in agreement, the output $\F_{r+1}$ is also in agreement.
\end{proof}

@@

\begin{lemma}
\label{lemma:fairness}
If for an arbitrary round $r$,
$\F_r$ is known by all correct nodes then
every node with a CP block in $\C_{r+1}$ should have an equal probability to be elected as a facilitator in $\F_{r+1}$.
\end{lemma}
\begin{proof}
We already established that $|\C_{r+1}| \ge N - t \ge n$ from \Cref{lemma:validity}.
Then the proof directly follows from the random oracle model.
Recall that the luck value is computed using $\textsf{H}(\C_{r+1} || pk_u)$.
Since $pk_u$ is unique for every node that has a CP block in $\C_{r+1}$,
the output of $\textsf{H}(\cdot)$ is uniformly random.
As a result, the output of $\textsf{get\_facilitators}(\C_{r+1})$ is a random permutation of all the nodes in $\C_{r+1}$ truncated to $n$.
Thus, every node has the same probability of becoming a facilitator.
\end{proof}

@@

$\Pi_v$ directly uses \Cref{alg:get-validity}, which is also the validity definition,
so the validity property is also satisfied.
\end{proof}

\Cref{theorem:validation-agreement} shows that agreement on CP blocks would lead to agreement on TX blocks when the nodes are running the validation protocol.
Like in our prior work~\cite{implicitconsensus}, we call this behaviour implicit consensus.
One of the main advantages of our scheme over running a consensus algorithm on all the transactions is that 
the transaction throughput is no longer dependent on the consensus algorithm---ACS.
This enables horizontal scalability where adding new nodes would lead to higher global transaction rate.
In addition, a convenient consequence of \Cref{theorem:validation-agreement} is unforgeability.
That is, no adversary can create two chains,
$F = \{ \dots, t, c\}$ and $F' = \{ \dots, t', c\}$,
with correct hash pointers and the same end of chain $c$ with non-negligible probability.
% Another way to think about it is if a transaction can be forged, then it cannot be agreed.

@@

\begin{lemma}
There exists a valid transaction that cannot be eventually validated.
\end{lemma}
\begin{proof}
Suppose nodes $u$ and $v$ correctly performed the TX protocol to create a transaction.
Then when $u$ wants to validate it, it does so by sending \texttt{vd\_req} message to $v$.
$v$ can act maliciously and ignore all \texttt{vd\_req} message from $u$, and then the transaction can never be validated.
\end{proof}
Although this is a negative result, it does not put the adversary in an advantageous position.
If the adversary is observed to ignore validation requests, then the honest nodes may prefer not to transact with her in the future.
Thus, to stay relevant in the system, the adversary needs to comply with the protocol.

@@

\section{Performance analysis}
\label{sec:performance-analysis}
This section aims to answer whether our system is horizontally scalable analytically.
We begin by looking at the communication and time complexity of the consensus protocol $\Pi_c$,
and then the bandwidth requirement for a single transaction.
We build on top of those results to analyse the global throughput.

@@

\subsection{Duration of the consensus protocol}
\label{sec:cons-duration}
To make arguments on the duration of $\Pi_c$, the bandwidth or throughput in our purely asynchronous model,
which are concepts that depend on time,
we must make additional assumptions.
Note that the duration we are interested in is not the same as the time complexity typically used in distributed systems.
In the analysis of distributed algorithms, time complexity is often in terms of the number of rounds.
For example, ACS runs in a constant number of rounds because of its sub-protocols---reliable broadcast and binary Byzantine consensus---also run in a constant number of rounds~\cite{miller2016honey}.
However, in practice, making a unit of communication always has some overhead associated with it, for example serialising and writing it to some network socket.
Hence, for the remainder of our performance analysis, we add the following to our computational model.
For every unit of communication, we assume they take some non-negligible but constant time to perform.
Hence, from our results in~\Cref{sec:cons-complexity} and the fact that $\Pi_c$ runs in a constant number of rounds,
it follows that $\Pi_c$ has a duration of $O(N)$.

\subsection{Communication cost for transactions}
\label{sec:communication-cost-for-tx}
Geared with the results above,
we are ready to analyse the amount of data required to be transmitted over a link per transaction,
which we call the communication cost per transaction.
To create and then validate a transaction,
the communication cost per transaction is of $O(l)$,
where $l$ is the length of the agreed fragment.
This result can be seen from the fact that the largest message by far is the \texttt{vd\_resp} message,
which contains the agreed fragment.
The other messages (\texttt{tx\_req}, \texttt{tx\_resp} and \texttt{vd\_req}) are constant factors.
If we assume that every node performs transactions at a constant rate of $r_{\text{tx}}$ per second, then
$$l = r_{\text{tx}} D_{\text{c}},$$
where $D_{\text{c}}$ is the duration of a round of the consensus protocol.
But from \Cref{sec:cons-duration},
we know that $D_{\text{c}}$ is of $O(N)$; thus the communication cost per transaction is $O(N)$.
This result is intuitive because round duration would be longer if there are more CP blocks (more $N$), which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
This behaviour is verified experimentally in \Cref{ch:implementation}.

\subsection{Linear global throughput}
\label{sec:global-throughput}
Using our results so far, we can analyse the global throughput.
First, we clarify the bandwidth definition, which is ``the data rate at which a network link or a network path can transfer''~\cite{prasad2003bandwidth}.
Now, suppose every node has $N$ links which connect to every other node,
every link has a fix bandwidth $C$ and every node makes $r_{\text{tx}}$ transactions per second.
Then we have the inequality
$$NC \ge r_{\text{tx}} l,$$
where $l$ is the length of the agreed fragment as before.
The inequality suggests that the rate for which transactions and validations are made cannot exceed the total bandwidth of all the links.

We note that the inequality does not hold if the node is only transacting with a subset of the population,
which is because it cannot use all the bandwidth available in all the links.
In the extreme case, if the node is only transacting with one other node, then it can only use the bandwidth of one link which is only $C$.
However, if that is the case, we can intelligently cache the \texttt{vd\_resp} messages as described in~\Cref{sec:caching}.
Hence, we analyse the worst case where every node transacts with a random node from the population,
and a new \texttt{vd\_resp} must be sent for every \texttt{vd\_req} message.

Consider the case where the system is making use of all the bandwidth, i.e. $NC = r_{\text{tx}} l$.
Recall that $l$ is of $O(N)$, which means LHS and RHS both grow linearly with respect to $N$.
Hence, there exists some $r_\text{tx}$ that makes use of all the available bandwidth regardless of $N$.
Finally, if every node in the network is transacting at $r_{\text{tx}}$ per second,
then the global throughput (in terms of transactions per second) is linear w.r.t. the population size $N$.
We also maintain a constant transaction rate by the same argument if the system is not making use of all the bandwidth.
Both claims are verified experimentally in~\Cref{ch:implementation}.

### 2017-08-22 thesis/chapters/ch5-implementation.tex
Thanks Johan
Edit grammar etc.

We evaluate our system experimentally to ultimately understand its scalability characteristics as well as its limitations.
Concretely,
we give an open source prototype implementation that runs on a network with over 1000 node.
Nodes in our implementation make transactions autonomously with each other and then we collect the results for analysis.

@@

The structure of the implementation is primarily made up of three modules---\texttt{acs}, \texttt{trustchain} and \texttt{node}.
\texttt{acs}, as its name suggests, implements ACS.
ACS uses erasure code in one of its sub-protocols (reliable broadcast).
Thus we use the liberasurecode\footnote{\url{https://github.com/openstack/liberasurecode}} library for its Reed-Solomon error correcting code functionality.
An implementation detail is that liberasurecode cannot create more than 32 code blocks, which is hardcoded in the source file\footnote{\url{https://github.com/openstack/liberasurecode/blob/0794b31c623e4cede76d66be730719d24debcca9/include/erasurecode/erasurecode.h\#L35}
}, we discuss the effect of this in~\Cref{sec:experimental-setup}.
The \texttt{acs} module provides a small interface to the caller to start and stop the consensus process and also retrieve results.
The \texttt{trustchain} module implements the Extended TrustChain data structure.
It also provides the essential algorithm necessary to interact with Extended TrustChain such as 
$\textsf{new\_tx}(\cdot)$, $\textsf{new\_cp}(\cdot)$ and $\textsf{agreed\_fragment}(\cdot)$.
Finally, the \texttt{node} module ties everything together.
It implements the consensus protocol, the transaction protocol and the validation protocol.

Every node keeps a persistent TCP connection with every other node.
This creates a fully connected network for our experiment.
It is certainly not ideal in real world scenarios where nodes may have limited resources (e.g. sockets).
But as a prototype, it is sufficient to run a network of over a thousand nodes and experiment with it.

@@

The goal of the experiment is to run \textsc{Checo} and analyse the communication cost, the consensus duration and the global throughput.
We investigate these properties under the following parameters.
\begin{enumerate}
  \item The number of facilitators $n \in \{4, 8, \dots, 32\}$.
        The maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation},
        but our results in~\Cref{sec:varying-facilitator-experiment} give a good indication of how our system may function for a larger $n$.
  \item The population size $N \in \{200, 300, \dots, 1200\}$.
        $N$ stops at 1200 is due to our physical setup, which we describe next.
  \item The two modes of transaction.
        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with one predetermined node. 
        This minimises the data volume per validated transaction because agreed fragment can be cached.
        The second mode is in the other extreme which we call the \emph{random-neighbour} mode,
        where every transaction is with a random node out of the $N$ nodes in the system.
        It is unlikely that the agreed fragments can be reused.
\end{enumerate}

The experiment is run on the DAS-5\footnote{\url{https://www.cs.vu.nl/das5/}}.
From now on, we use ``machines'' to refer to DAS-5 nodes and ``nodes'' to refer to a running instance in our system.
On DAS-5 we use up to 30 machines, for each machine we use up to 40 nodes.
This gives us the aforementioned 1200 number.
With this setup, we cannot run more nodes because the every machine only has 65535 ports available (minus the reserved ones).
But 40 nodes each need 1200 TCP connections which is 48000 TCP connections per machine and that is inching close to the limit.
While it is possible to have more TCP connections per machine,
but it requires additional network interface which is something we do not control on the DAS-5.
Nevertheless, running the system with 1200 nodes gives a good indication of its scalability properties as we show later.

@@

\Cref{fig:round-comms} shows the relationship between the communication cost of the consensus protocol per round and the population size.
The most important aspect is that these results show a linear increase.
This reinforces our analytical result in~\Cref{sec:cons-complexity}.
Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
the magnitude of the communication cost is similar.
Both peak at about 100 MB,
which is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
Finally, the rate for which the communication cost increases is higher when the number of facilitators is higher,
which is also expected because the ACS algorithm has a polynomial term with respect to $n$ in its communication complexity.

We are also interested in how communication costs translate to time.
Hence, for the same experiment, we record the duration in seconds and the result is shown in~\Cref{fig:round-duration}.
Interestingly, the duration is not entirely linear.
We attribute this behaviour to the extra time needed to hash the CP blocks in the consensus result to compute the luck value.
Since if $N$ increases, every node must also perform more hash operations.
These results do not conform to analytical result in~\Cref{sec:cons-duration}.
Nevertheless, the difference is minor and there are ways to optimise the luck value computation.
For example, the luck value can be computed by the facilitators and are sent with the consensus result.
Then the non-facilitator nodes simply use it if there are enough signatures.

@@

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/tx-communication-cost-vs-population}
    \caption{Transactions are with fixed neighbours.}
    \label{fig:tx-comms-fixed}
  \end{subfigure}%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/tx-communication-cost-vs-population}
    \caption{Transactions are with random neighbours.}
    \label{fig:tx-comms-random}
  \end{subfigure}
  }
  \caption{Experimental result of the communication cost per verified transaction.
  It has a similar, near linear, trend as~\Cref{fig:round-duration}.
  Fluctuation for the fixed-neighbour mode exists because the cache mechanism is unpredictable.}
  \label{fig:tx-comms}
\end{figure}

We argued that the communication cost per verified transaction is of $O(N)$ (\Cref{sec:communication-cost-for-tx}).
To verify the argument, we plot the relationship between the communication cost of for every validated transaction and population size in~\Cref{fig:tx-comms}.
We observe a near linear relationship, which is due to the near linear relationship of the communication duration mentioned before.
Again, we believe the difference is minor and it is possible to remove the extra overhead.

More interestingly, there is a large difference in communication cost between the two modes of transaction.
When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round.
On the other hand,
if every node is transacting with a random node, then it is likely the case that one agreed fragment needs to be communicated for every transaction.
Hence the communication cost we see in~\Cref{fig:tx-comms-random} is much higher than in~\Cref{fig:tx-comms-fixed}.

Some fluctuations exist in~\Cref{fig:tx-comms-fixed}, this is due to our caching mechanism.
We send validation requests at the same rate as transactions.
Upon receiving a (remote) agreed fragment, the caching mechanism inspects all the transactions in the agreed fragment and attempts to verify as many as it can,
rather than just the transaction in the original validation request.
However, it may be the case that the agreed fragments arrive later than the validation request interval.
Then it is possible to have sent two or more validation requests for some transactions in the same local agreed fragment.
In this case, the remote would respond with two or more of the same agreed fragments,
which results in extra (wasted) communication cost, and this is the source of the fluctuation seen in~\Cref{fig:tx-comms-fixed}.
The result in~\Cref{fig:tx-comms-random} reinforces our argument.
It is a lot more stable because for every transaction it is almost always the case that a new agreed fragment is needed to validate it.

@@

The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is caused by the caching mechanism mentioned earlier.
If a new agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
The low transaction rate in~\Cref{fig:global-throughput-random} is caused by the fact that the network infrastructure cannot keep up with our demand.
In practice, we do not expect such behaviour to occur as it is possible to cache agreed fragments.

@@

\section{Communication cost with varying number of facilitators}
\label{sec:varying-facilitator-experiment}
Up to this point we focused on the effect of communication cost, throughput and so on with respect to the population size.
In this section, we consider a varying number of facilitators,
which gives us an insight into our system performance may perform when the number of facilitators is higher than 32.

@@

\Cref{fig:facilitators-round-comms} shows the communication cost of the consensus protocol for varying numbers of facilitators.
There is strong evidence that the communication cost grows polynomially.
We expect this behaviour because of there are polynomial terms in the ACS communication complexity---$O(n^2|v| + \lambda n^3 \log n)$.
This also means that the rounds would take longer to complete and transactions would take longer to verify.
On the other hand, we get better fault tolerance.
Thus a linear increase in fault tolerance ($t$) costs a polynomial increase in communication cost.

>>>>

We evaluate our system experimentally to understand its scalability characteristics as well as its limitations.
Concretely,
we give a free and open-source prototype implementation that runs on a network with over 1000 node.
Nodes in our implementation make transactions autonomously with each other, and then we collect the results for analysis.

@@

The structure of the implementation is primarily made up of three modules---\texttt{acs}, \texttt{trustchain} and \texttt{node}.
\texttt{acs}, as its name suggests, implements ACS.
ACS uses erasure code in one of its sub-protocols (reliable broadcast).
Thus we use the liberasurecode\footnote{\url{https://github.com/openstack/liberasurecode}} library for its Reed-Solomon error correcting code functionality.
An implementation detail is that liberasurecode cannot create more than 32 code blocks, which is hard-coded in the source file\footnote{\url{https://github.com/openstack/liberasurecode/blob/0794b31c623e4cede76d66be730719d24debcca9/include/erasurecode/erasurecode.h\#L35}
}, we discuss the effect of this in~\Cref{sec:experimental-setup}.
The \texttt{acs} module provides a small interface to the caller to start and stop the consensus process and also retrieve results.
The \texttt{trustchain} module implements the Extended TrustChain data structure.
It also provides the essential algorithm necessary to interact with Extended TrustChain such as 
$\textsf{new\_tx}(\cdot)$, $\textsf{new\_cp}(\cdot)$ and $\textsf{agreed\_fragment}(\cdot)$.
Finally, the \texttt{node} module ties everything together.
It implements the consensus protocol, the transaction protocol and the validation protocol.

Every node keeps a persistent TCP connection with every other node,
which creates a fully connected network for our experiment.
It is certainly not ideal in real world scenarios where nodes may have limited resources (e.g. sockets).
But as a prototype, it is sufficient to run a network of over a thousand nodes and experiment with it.

@@

The goal of the experiment is to run \textsc{Checo} and analyse the communication cost, the consensus duration and the global throughput.
We investigate these properties under the following parameters.
\begin{enumerate}
  \item The number of facilitators $n \in \{4, 8, \dots, 32\}$.
        The maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation},
        but our results in~\Cref{sec:varying-facilitator-experiment} give a good indication of how our system may function for a larger $n$.
  \item The population size $N \in \{200, 300, \dots, 1200\}$.
        The reason why $N$ only goes up to 1200 is due to our physical setup, which we describe below.
  \item The two modes of transaction.
        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with one predetermined node. 
        It minimises the data volume per validated transaction because agreed fragment can be cached.
        The second mode is in the other extreme which we call the \emph{random-neighbour} mode,
        where every transaction is performed with a random node out of the $N$ nodes in the system.
        It is unlikely that the agreed fragments can be reused.
\end{enumerate}

The experiment is run on the DAS-5\footnote{\url{https://www.cs.vu.nl/das5/}}.
From now on, we use ``machines'' to refer to DAS-5 nodes and ``nodes'' to refer to a running instance in our system.
On DAS-5 we use up to 30 machines, for each machine we use up to 40 nodes,
which gives us the aforementioned 1200 number.
With this setup, we cannot run more nodes because the every machine only has 65535 ports available (minus the reserved ones).
But 40 nodes each need 1200 TCP connections which is 48000 TCP connections per machine and that is inching close to the limit.
While it is possible to have more TCP connections per machine,
but it requires additional network interface which is something we do not control on the DAS-5.
Nevertheless, running the system with 1200 nodes gives a good indication of its scalability properties as we show later.

@@

\Cref{fig:round-comms} shows the relationship between the communication cost of the consensus protocol per round and the population size.
The most important aspect is that these results show a linear increase,
which reinforces our analytical result in~\Cref{sec:cons-complexity}.
Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
the magnitude of the communication cost is similar.
Both peak at about 100 MB,
which is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
Finally, the rate for which the communication cost increases is higher when the number of facilitators is higher,
which is also expected because the ACS algorithm has a polynomial term w.r.t. $n$ in its communication complexity.

We are also interested in how communication costs translate to time.
Hence, for the same experiment, we record the duration in seconds and the result is shown in~\Cref{fig:round-duration}.
Interestingly, the duration is not entirely linear.
We attribute this behaviour to the extra time needed to hash the CP blocks in the consensus result to compute the luck value.
Since if $N$ increases, every node must also perform more hash operations.
These results do not conform to analytical result in~\Cref{sec:cons-duration}.
Nevertheless, the difference is minor, and there are ways to optimise the luck value computation.
For example, the luck value can be computed by the facilitators and are sent with the consensus result.
Then the non-facilitator nodes simply use it if there are enough signatures.

@@

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/tx-communication-cost-vs-population}
    \caption{Transactions are with fixed neighbours.}
    \label{fig:tx-comms-fixed}
  \end{subfigure}%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/tx-communication-cost-vs-population}
    \caption{Transactions are with random neighbours.}
    \label{fig:tx-comms-random}
  \end{subfigure}
  }
  \caption{Experimental result of the communication cost per verified transaction.
  It has a similar, nearly linear, trend as~\Cref{fig:round-duration}.
  Fluctuation for the fixed-neighbour mode exists because the cache mechanism is unpredictable.}
  \label{fig:tx-comms}
\end{figure}

We argued that the communication cost per verified transaction is of $O(N)$ (\Cref{sec:communication-cost-for-tx}).
To verify the argument, we plot the relationship between the communication cost of for every validated transaction and population size in~\Cref{fig:tx-comms}.
We observe a nearly linear relationship, which is due to the nearly linear relationship of the communication duration mentioned before.
Again, we believe the difference is minor, and it is possible to remove the extra overhead.

More interestingly, there is a large difference in communication cost between the two modes of transaction.
When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round to validate all transactions of that round.
On the other hand,
if every node is transacting with a random node, then it is likely the case that one agreed fragment needs to be communicated for every transaction.
Hence the communication cost we see in~\Cref{fig:tx-comms-random} is much higher than in~\Cref{fig:tx-comms-fixed}.

Some fluctuations exist in~\Cref{fig:tx-comms-fixed}, which is due to our caching mechanism.
We send validation requests at the same rate as transactions.
Upon receiving a (remote) agreed fragment, the caching mechanism inspects all the transactions in the agreed fragment and attempts to verify as many as it can,
rather than just the transaction in the original validation request.
However, it may be the case that the agreed fragments arrive later than the validation request interval.
Then it is possible to have sent two or more validation requests for some transactions in the same local agreed fragment.
In this case, the remote would respond with two or more of the same agreed fragments,
which results in extra (wasted) communication cost, and this is the source of the fluctuation seen in~\Cref{fig:tx-comms-fixed}.
The result in~\Cref{fig:tx-comms-random} reinforces our argument.
It is a lot more stable because for every transaction it is almost always the case that a new agreed fragment is needed to validate it.

@@

The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is caused by the caching mechanism mentioned earlier.
If a new agreed fragment needs to be transmitted to validate every transaction,
then it puts a toll on the network infrastructure.
The low transaction rate in~\Cref{fig:global-throughput-random} is caused by the fact that the network infrastructure cannot keep up with our demand.
In practice, we do not expect such behaviour to occur as it is possible to cache agreed fragments.

@@

\section{Communication cost with varying number of facilitators}
\label{sec:varying-facilitator-experiment}
Up to this point, we focused on the effect of communication cost, throughput and so on with respect to the population size.
In this section, we consider a varying number of facilitators,
which gives us an insight into our system performance may perform when the number of facilitators is higher than 32.

@@

\Cref{fig:facilitators-round-comms} shows the communication cost of the consensus protocol for varying numbers of facilitators.
There is substantial evidence that the communication cost grows polynomially.
We expect this behaviour because of there are polynomial terms in the ACS communication complexity---$O(n^2|v| + \lambda n^3 \log n)$,
which means that the rounds would take longer to complete and transactions would take longer to verify.
On the other hand, we get better fault tolerance.
Thus a linear increase in fault tolerance ($t$) costs a polynomial increase in communication cost.
